---
title: "Summer2023UseAfterFire"
output: html_document
date: "2023-09-26"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#### Libraries ####

library(janitor)
library(tidyverse)
library(ggthemes)
library(lubridate)
library(lme4)
library(readxl)
library(writexl)
library(sf)
library(MuMIn)
library(scales)
library(sjPlot)
library(rsq)
library(betareg)
library(broom)
library(caret)
library(janitor)
library(performance)
library(gam)
library(mgcv)
theme_set(theme_minimal())
```

Importing data sets
```{r}
#### Importing veg data  and locations ####

#Veg data
veg_data_2023 <- read_xlsx("/Volumes/Samsung_T5/BOBWHITE_DATA/Clean/VegSurveys/VegSurveys2023.xlsx") %>%
  clean_names() %>%
  mutate(date = as.Date(date)) 

#Lets seperated out veg data to isolate veg plots burned 17 April
veg_data_2023_units1_5 <- veg_data_2023 %>%
  dplyr::filter(unit %in% 1:5)

veg_data_2023_unit6 <- veg_data_2023 %>%
  dplyr::filter(unit %in% 6)


#Importing all summer locations. This includes burn status, which we got from ArcGIS
s23_locs <- read_xlsx("/Volumes/Samsung_T5/BOBWHITE_DATA/Clean/SummerLocations/Summer2023/2023_burn_locs_burned_unburned_CORRECTED.xlsx") %>%
  clean_names() %>%
  mutate(date = ymd(date)) %>%
  dplyr::filter(status != "D") %>%
  mutate(burned = as.double(burned))

#Importing a spreadsheet of all locations from burn unit burned in 17 April (these rows still in s23_locs... this is a subset).
#These points are already properly identified as burned/unburned based on burn date. We will need to delete this subset from s23_locs. And add veg data and dsf to each data set before joining.
s23_april17subset <- read_xlsx("/Volumes/Samsung_T5/BOBWHITE_DATA/Clean/SummerLocations/Summer2023/s23LocsIn17AprilBurnUnit.xlsx") %>%
  clean_names() %>%
  mutate(date = as.Date(date)) %>%
  dplyr::filter(date > "2023-04-17",
                status != "D") 


#Here, we delete the subset from s23 locs. We will update s23 locs with veg data and dsf. Update the subset with same info, and then join them.
#Now, we must replace rows in s23_locs with identical rows in the subset (based on band_numb, date, northing, and easting)
s23_locs_main <- rows_delete(s23_locs, s23_april17subset, by = c("band_numb", "date", "northing", "easting"))


#Adding DSF to main dataset. Most locations burned March 21 2023. We'll add the subset of other locations next.
burn_date_2023_most <- ymd("2023-03-21")

s23_locs_main$dsf <- as.numeric(s23_locs_main$date - burn_date_2023_most)


#Now we add DSF to subset.
burn_date_subset <- ymd("2023-04-17")


s23_april17subset$dsf <- as.numeric(s23_april17subset$date - burn_date_subset)

```

Summarizing veg data and adding to to s23_locs_main
```{r}
#### Adding veg data to s23 locs ####]

#Summarizing veg data.
veg_data_2023_units1_5_summary <- veg_data_2023_units1_5 %>%
  group_by(date) %>%
  summarize(hg = mean(max_grasses_cm),
            hf = mean(max_forbs_cm),
            hs = mean(max_shrubs_cm),
            pg = mean(percent_grasses),
            pf = mean(percent_forbs),
            ps = mean(percent_shrubs))

#adding week column 
veg_data_2023_units1_5_summary$week <- week(veg_data_2023_units1_5_summary$date)

#We'll need to update this manually. Some done on sat, some on sun.
#writexl::write_xlsx(veg_data_2023_units1_5_summary, "/Users/jeffgrayum/Downloads/vegsummary_23_1_5.xlsx")

#Reading back in after correcting.
veg_data_2023_units1_5_summary <- read_xlsx("/Volumes/Samsung_T5/BOBWHITE_DATA/Clean/VegSurveys/vegsummary_23_1_5.xlsx")

#Now we will add week column to summer locs (main)
s23_locs_main$week <- week(s23_locs_main$date)

#Now, we join based on week.
s23_locs_main_veg <- inner_join(s23_locs_main, veg_data_2023_units1_5_summary, by = "week")

#taking a look
s23_locs_main_veg%>%
  view()

#renaming some columns
s23_locs_main_veg <- s23_locs_main_veg %>%
  rename(date = date.x,
         start_date_of_week = date.y) %>%
  select(-column1)

```

Summarizing veg data and adding to to s23_locs_subset
```{r}
#### Adding veg data to s23 locs ####]

#Summarizing veg data.
veg_data_2023_unit6_summary <- veg_data_2023_unit6 %>%
  group_by(date) %>%
  summarize(hg = mean(max_grasses_cm),
            hf = mean(max_forbs_cm),
            hs = mean(max_shrubs_cm),
            pg = mean(percent_grasses),
            pf = mean(percent_forbs),
            ps = mean(percent_shrubs))

#adding week column 
veg_data_2023_unit6_summary$week <- week(veg_data_2023_unit6_summary$date)

#We'll need to update this manually. Some done on sat, some on sun.
#writexl::write_xlsx(veg_data_2023_unit6_summary, "/Users/jeffgrayum/Downloads/vegsummary_23_6.xlsx")

#Reading back in after correcting.
veg_data_2023_unit6_summary <- read_xlsx("/Volumes/Samsung_T5/BOBWHITE_DATA/Clean/VegSurveys/vegsummary_23_6.xlsx")

#Now we will add week column to summer locs (main)
s23_april17subset$week <- week(s23_april17subset$date)

#writexl::write_xlsx(s23_april17subset, "/Users/jeffgrayum/Downloads/s23_april17subset.xlsx")

#Now, we join based on week.
s23_april17subset_veg <- inner_join(s23_april17subset, veg_data_2023_unit6_summary, by = "week")

#writexl::write_xlsx(s23_april17subset_veg, "/Users/jeffgrayum/Downloads/s23_april17subset_veg.xlsx")

#Required some cleaning, reading it back in:
s23_april17subset_veg_fixed <- read_xlsx("/Volumes/Samsung_T5/BOBWHITE_DATA/Clean/SummerLocations/Summer2023/s23_april17subset_veg_fixed.xlsx")

#taking a look
s23_april17subset_veg_fixed %>%
  view()

#renaming some columns
s23_april17subset_veg_fixed <- s23_april17subset_veg_fixed %>%
  rename(start_date_of_week = date.y)

```

Now, we can bind main data set and subset!
```{r}
s23_locs_veg_all_fixed <- rbind(s23_april17subset_veg_fixed, s23_locs_main_veg)

s23_locs_veg_all_fixed <- s23_locs_veg_all_fixed %>%
  arrange(date)

s23_locs_veg_all_fixed <- s23_locs_veg_all_fixed %>%
  dplyr::filter(date < "2023-08-05")
```


Adding a distance to edge column to locations
```{r}
#### Adding dte ####

#Importing map of burned areas, summer 2023.
#Reading burn compartments in again, but as an sf object
burn_compartments_2023 <- st_read("/Volumes/Samsung_T5/BOBWHITE_DATA/Maps/2023_burn_compartments/burn 2023 map/NoAG_NoHW/2023_burn_units_NoAg_NoHW.shp")

# Converting burn_compartments_2022 polygons to line geometries that represent their boundaries. This is necessary to calc min distances later.
burn_compartment_boundaries <- st_boundary(burn_compartments_2023)

# Converting tibble of points to an sf object
s23_locs_sf <- st_as_sf(s23_locs_veg_all_fixed, coords = c("easting", "northing"), crs = st_crs(burn_compartments_2023))

# Calculating the distance from each point to the nearest line boundary of the burn compartments, storing it in matrix
distance_matrix <- st_distance(s23_locs_sf, burn_compartment_boundaries)

# Applying the function min() to each row (1 does this -- we'd use 2 for columns) in the distance matrix
min_distances <- apply(distance_matrix, 1, min)

# Add the minimum distance to the original tibble as a new column
s23_locs_veg_all_fixed$distance_to_edge <- min_distances

write_xlsx(s23_locs_veg_all_fixed, "/Volumes/Samsung_T5/BOBWHITE_DATA/Clean/SummerLocations/Summer2023/s23_locs_veg_all_fixed.xlsx")
```



We're gonna try joining in R...
```{r}
km23 <- read_xlsx("/Users/jeffgrayum/Downloads/km23df.xlsx") %>%
  mutate(date = ymd(date)) %>%
  dplyr::filter(date >= "2023-03-22")

s23_locs_veg_birds_online <- s23_locs_veg %>%
  clean_names() %>%
  left_join(km23 %>% dplyr::select(n.risk, date), 
            by = "date")

write_xlsx(s23_locs_veg_birds_online, "/Users/jeffgrayum/Downloads/s23_locs_veg_birds_online.xlsx")
```


Making some distance to edge plots (Still, not the expected trend)
```{r}
#### Plotting DTE ####
s23_locs_veg_all_fixed %>%
  dplyr::filter(burned == 1) %>%
  ggplot(aes(dsf, distance_to_edge)) +
  geom_point() +
  geom_smooth() +
  labs(x = "DSF",
       y = "Distance to edge (m)",
       title = "Distance to edge following prescribed fire (2023)",
       subtitle = "Only looking at birds within burned units")

s23_locs_veg %>%
  dplyr::filter(burned == 1) %>%
  summarize(mean_dte = mean(distance_to_edge)) %>%
  print()

s23_locs_veg %>%
  dplyr::filter(burned == 0) %>%
  ggplot(aes(dsf, distance_to_edge)) +
  geom_point() +
  geom_smooth() +
  labs(x = "DSF",
       y = "Distance to edge (m)",
       title = "Distance to edge following prescribed fire (2023)",
       subtitle = "Only looking at birds within unburned units")

s23_locs_veg %>%
  dplyr::filter(burned == 0) %>%
  summarize(mean_dte = mean(distance_to_edge)) %>%
  print()
```

Now, we will determine the proportion of birds in burned areas each day.
```{r}
s23_locs_veg_birds_online_fixed_veg <- read_xlsx("/Volumes/Samsung_T5/BOBWHITE_DATA/Clean/SummerLocations/Summer2023/s23_locs_veg_birds_online_fixedVEG.xlsx") %>%
  mutate(date = as.Date())

# Counting unique birds signaled within a burned area each day
birds_in_burned_areas_each_day23 <- s23_locs_veg_birds_online_fixed_veg %>%
  dplyr::filter(burned == 1) %>%
  group_by(date) %>%
  summarize(birds_in_burned_area_each_day23 = n_distinct(band_numb))

# Joining into summer locs data
s23_locs_veg_birds_online_fixed_veg <- s23_locs_veg_birds_online_fixed_veg %>%
  left_join(birds_in_burned_areas_each_day23, by = "date")

# Handling NA values for birds_in_burned_area_each_day23
s23_locs_veg_birds_online_fixed_veg$birds_in_burned_area_each_day23[is.na(s23_locs_veg_birds_online_fixed_veg$birds_in_burned_area_each_day23)] <- 0

s23_locs_veg_birds_online_fixed_veg <- s23_locs_veg_birds_online_fixed_veg %>%
  rename(birds_in_burned_daily = birds_in_burned_area_each_day23)

# Calculate percentage of birds in burned areas
s23_locs_veg_birds_online_fixed_veg <- s23_locs_veg_birds_online_fixed_veg %>%
  mutate(percent_in_burned = (birds_in_burned_daily / birds_online))

# Handling NA values for percent_in_burned
s23_locs_veg_birds_online_fixed_veg$percent_in_burned[is.na(s23_locs_veg_birds_online_fixed_veg$percent_in_burned)] <- 0

#shortening name of df
s23rx_FV <- s23_locs_veg_birds_online_fixed_veg

writexl::write_xlsx(s23rx_FV, "/Users/jeffgrayum/Downloads/s23rx_FV.xlsx")
```

Plotting use over dsf
```{r}
s23rx_FV %>%
  ggplot(aes(dsf, percent_in_burned)) +
  geom_point()
```

