---
title: "Burned_Unburned_Locs_VegData"
output: html_document
date: "2023-09-11"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(janitor)
library(ggthemes)
library(lubridate)
library(lme4)
library(readxl)
library(writexl)
library(sf)
library(MuMIn)
theme_set(theme_minimal())
```

Importing datasets
```{r}
veg_data_2022 <- read_xlsx("/Volumes/Samsung_T5/BOBWHITE_DATA/Clean/VegSurveys2022.xlsx") %>%
  clean_names() %>%
  mutate(date = as.Date(date))

summer_locs <- read_xlsx("/Volumes/Samsung_T5/BOBWHITE_DATA/Clean/summer_2022_locs_burned_and_unburned_clean.xlsx") %>%
  mutate(date = as.Date(date)) %>%
  arrange(date) %>%
  as_tibble() %>%
  view()
```

First, lets summarize our veg data so we have average values for each date, rather than 50 samples.
```{r}
veg_data_2022_summary <- veg_data_2022 %>%
  group_by(date) %>%
  summarize(median_percent_grasses = median(percent_grasses),
            median_percent_forbs = median(percent_forbs),
            median_percent_shrubs = median(percent_shrubs),
            median_max_grasses = median(max_grasses_cm),
            median_max_forbs = median(max_forbs_cm),
            median_max_shrubs = median(max_shrubs_cm)) %>%
  as_tibble()

view(veg_data_2022_summary)
```

Lets add a distance to edge column for our points (Not sure if locations outside of burned areas should be negative?)
```{r}
#Importing map of burned areas, summer 2022.
#Reading burn compartments in again, but as an sf object
burn_compartments_2022 <- st_read("/Volumes/Samsung_T5/BOBWHITE_DATA/Maps/2022_burn_compartments/Polygons/2022_BurnedUnits_minusAG2.shp")

# Converting burn_compartments_2022 polygons to line geometries that represent their boundaries. This is necessary to calc min distances later.
burn_compartment_boundaries <- st_boundary(burn_compartments_2022)

# Converting tibble of points to an sf object
summer_locs_sf <- st_as_sf(summer_locs, coords = c("easting", "northing"), crs = st_crs(burn_compartments_2022))

# Calculating the distance from each point to the nearest line boundary of the burn compartments, storing it in matrix
distance_matrix <- st_distance(summer_locs_sf, burn_compartment_boundaries)

# Applying the function min() to each row (1 does this -- we'd use 2 for columns) in the distance matrix
min_distances <- apply(distance_matrix, 1, min)

# Add the minimum distance to the original tibble as a new column
summer_locs$distance_to_edge <- min_distances

#Printing
print(summer_locs)

#This looks right, and similar to what we got in other analysis.
summer_locs %>%
  dplyr::filter(burned == 1) %>%
  ggplot(aes(DSF, distance_to_edge)) +
  geom_point() +
  geom_smooth() +
  labs(x = "DSF",
       y = "Distance to edge (m)",
       title = "Distance to edge following prescribed fire",
       subtitle = "Only looking at birds within burned units")

summer_locs %>%
  dplyr::filter(burned == 1) %>%
  summarize(mean_distance = mean(distance_to_edge))


summer_locs %>%
  dplyr::filter(burned == 1) %>%
  ggplot(aes(distance_to_edge)) +
  geom_histogram(binwidth = 10, fill = "midnight blue", alpha = 0.8) + 
  geom_vline(xintercept = 27.8, linetype = "dashed", color = "gray") +
  labs(x = "Distance to edge",
       y = "",
       title = "Distances to edge following prescribed fire",
       subtitle = "Only looking at birds within burn compartments")

  

summer_locs %>%
  dplyr::filter(burned == 0) %>%
  ggplot(aes(DSF, distance_to_edge)) +
  geom_point() +
  geom_smooth() +
  labs(x = "DSF",
       y = "Distance to edge",
       title = "Distance to edge of burned compartments following prescribed fire",
       subtitle = "Only looking at birds located outside of burned units")

summer_locs %>%
  dplyr::filter(burned == 0) %>%
  summarize(mean_distance = mean(distance_to_edge))


summer_locs %>%
  dplyr::filter(burned == 0) %>%
  ggplot(aes(distance_to_edge)) +
  geom_histogram(binwidth = 20, fill = "midnight blue", alpha = 0.8) + 
  geom_vline(xintercept = 96.9, linetype = "dashed", color = "gray") +
  labs(x = "Distance to edge",
       y = "",
       title = "Distances to edge following prescribed fire",
       subtitle = "Only looking at birds outside of burn compartments")

summer_locs %>%
  mutate(burned = factor(burned, labels = c("Unburned", "Burned"))) %>%
  ggplot(aes(x = burned, y = distance_to_edge)) +
  geom_boxplot(fill = "midnight blue", alpha = 0.5) +
  labs(x = "Burn Status",
       y = "Distance to Edge",
       title = "Distance to edge of burn compartments following prescribed fire")

summer_locs %>%
  mutate(burned = factor(burned, labels = c("Unburned", "Burned"))) %>%
  ggplot(aes(x = burned, y = distance_to_edge)) +
  geom_violin(fill = "midnight blue", alpha = 0.5) +
  labs(x = "Burn Status",
       y = "Distance to Edge",
       title = "Distance to edge following prescribed fire")

```

Lets add in age and sex info for our location data. This may be useful later.
```{r}
#The fate spreadsheet has age and sex info for each bird.
#Importing "fate" spreadsheet
fate <- rio::import("/Volumes/Samsung_T5/BOBWHITE_DATA/Clean/fate_clean_2021_22.xlsx", setclass = "tibble") %>%
  clean_names() %>%
  rename(band_numb = band_number) 

#Now we bind age and sex into locations, based on band_numb
summer_locs <- summer_locs %>%
  left_join(fate %>% 
              dplyr::select(band_numb, age, sex), 
            by = "band_numb")
```

Trying to join veg summary data with Summer 2022 locations.
We have locations almost everyday, but veg data only once a week. 
First, we will try adding a column "week" to join by (this looks like it worked).
```{r}
#Adding week column to each tibble
summer_locs$week <- week(summer_locs$date)
veg_data_2022_summary$week <- week(veg_data_2022_summary$date)

#I am surprised this worked.
summer_locs %>%
  view()
veg_data_2022_summary %>%
  view()

#Now, we join based on week.
summer_locs_veg_data <- inner_join(summer_locs, veg_data_2022_summary, by = "week")

#taking a look
summer_locs_veg_data %>%
  view()

summer_locs_veg_data <- summer_locs_veg_data%>%
  rename(date = date.x,
         start_date_of_week = date.y)

#write_xlsx(summer_locs_veg_data, "/Volumes/Samsung_T5/BOBWHITE_DATA/Clean/summer_locs_veg_dataFIXEDdistances.xlsx")
```

Trying to determine the proportion of bobwhite that were signaled in burned areas each day (birds in burned areas/unique band numbers signaled that week)
```{r}
# Counting birds online each week
birds_online_each_week <- summer_locs_veg_data %>%
  group_by(week) %>%
  summarize(birds_online = n_distinct(band_numb))

# Adding count to tibble
summer_locs_veg_data <- summer_locs_veg_data %>%
  left_join(birds_online_each_week, by = "week")

# Handling NA values for birds_online
summer_locs_veg_data$birds_online[is.na(summer_locs_veg_data$birds_online)] <- 0

# Counting unique birds signaled within a burned area each day
birds_in_burned_areas_each_day <- summer_locs_veg_data %>%
  dplyr::filter(burned == 1) %>%
  group_by(date) %>%
  summarize(birds_in_burned_area = n_distinct(band_numb))

# Joining into summer locs data
summer_locs_veg_data <- summer_locs_veg_data %>%
  left_join(birds_in_burned_areas_each_day, by = "date")

# Handling NA values for birds_in_burned_area
summer_locs_veg_data$birds_in_burned_area[is.na(summer_locs_veg_data$birds_in_burned_area)] <- 0

# Calculate percentage of birds in burned areas
summer_locs_veg_data_perc_burned <- summer_locs_veg_data %>%
  mutate(percent_in_burned = (birds_in_burned_area / birds_online) * 100)

# Handling NA values for percent_in_burned
summer_locs_veg_data_perc_burned$percent_in_burned[is.na(summer_locs_veg_data_perc_burned$percent_in_burned)] <- 0


ave_burned_per_week <- summer_locs_veg_data_perc_burned %>%
  group_by(week) %>%
  summarize(mean_per_burned = mean(percent_in_burned))

ave_burned_per_week <- ave_burned_per_week %>%
  mutate(week = (week - 15))

ave_burned_per_week %>%
  dplyr::filter(week <=15) %>%
  ggplot(aes(week, mean_per_burned)) +
  geom_line() +
  labs(x = "Weeks following prescribed fire",
       y = "Mean percentage of birds online that were located in burned areas",
       title = "Average percentage of birds online that were located in burned areas each week")


summer_locs_veg_data_perc_burned %>%
  ggplot(aes(DSF, percent_in_burned)) +
  geom_point() +
  geom_smooth()

```


```{r}
summer_locs_veg_data %>%
  dplyr::filter(burned == 1) %>%
  ggplot(aes(median_max_grasses, distance_to_edge)) +
  geom_point() +
  geom_smooth() +
  labs(x = "Median height of grasses",
       y = "Distance to edge",
       title = "Grass height vs Dist to Edge")

summer_locs_veg_data %>%
  dplyr::filter(burned == 1) %>%
  ggplot(aes(median_max_forbs, distance_to_edge)) +
  geom_point() +
  geom_smooth() +
  labs(x = "Median height of forbs",
       y = "Distance to edge",
       title = "Forb height vs Dist to Edge")

summer_locs_veg_data %>%
  dplyr::filter(burned == 1) %>%
  ggplot(aes(median_max_shrubs, distance_to_edge)) +
  geom_point() +
  geom_smooth() +
  labs(x = "Median height of shrubs",
       y = "Distance to edge",
       title = "Shrub height vs Dist to Edge")

summer_locs_veg_data %>%
  dplyr::filter(burned == 1) %>%
  ggplot(aes(median_percent_grasses, distance_to_edge)) +
  geom_point() +
  geom_smooth() +
  labs(x = "Median percent cover of grasses",
       y = "Distance to edge",
       title = "Percent cover grass vs Dist to Edge")

summer_locs_veg_data %>%
  dplyr::filter(burned == 1) %>%
  ggplot(aes(median_percent_forbs, distance_to_edge)) +
  geom_point() +
  geom_smooth() +
  labs(x = "Median percent cover of forbs",
       y = "Distance to edge",
       title = "Percent cover forbs vs Dist to Edge")

summer_locs_veg_data %>%
  dplyr::filter(burned == 1) %>%
  ggplot(aes(median_percent_shrubs, distance_to_edge)) +
  geom_point() +
  geom_smooth() +
  labs(x = "Median percent cover of shrubs",
       y = "Distance to edge",
       title = "Percent cover shrubs vs Dist to Edge")
```


Trying to run a mixed-model.
```{r}
#First I think we gotta standardize our covariates?
#Doing this in two steps.. one step for percent cover class, one for max height. Not sure if this is even necessary.
summer_locs_veg_data_modified <- summer_locs_veg_data %>% 
  mutate_at(vars(median_percent_grasses:median_max_shrubs,
                 DSF:distance_to_edge), .funs = function(x){as.numeric(scale(x, center = T))}) 

summer_locs_veg_data_modified %>%
  view()
  

#Starting by running a global model.
#Fitting glm
prob_burn_area_global <- glmer(burned ~ median_percent_grasses + median_percent_forbs + median_percent_shrubs + median_max_grasses + median_max_forbs + median_max_shrubs + DSF + (1|band_numb),
                 data = summer_locs_veg_data_modified,
                 na.action = "na.pass",
                 family = "binomial")
 
#Viewing summary 
summary(prob_burn_area_global) 

#This model failed to converge

#Now looking at only height
prob_burn_area_height <- glmer(burned ~ median_max_grasses + median_max_forbs + median_max_shrubs + (1|band_numb),
                 data = summer_locs_veg_data_modified,
                 na.action = "na.pass",
                 family = "binomial")
 
#Viewing summary 
summary(prob_burn_area_height) 

#Now looking at only height
prob_burn_area_veg_area <- glmer(burned ~ median_percent_grasses + median_percent_forbs + median_percent_shrubs + (1|band_numb),
                 data = summer_locs_veg_data_modified,
                 na.action = "na.pass",
                 family = "binomial")
 
#Viewing summary 
summary(prob_burn_area_veg_area) 

#Only looking at DSF
prob_burn_area_DSF <- glmer(burned ~ DSF + (1|band_numb),
                 data = summer_locs_veg_data_modified,
                 na.action = "na.pass",
                 family = "binomial")
 
#Viewing summary 
summary(prob_burn_area_DSF) 


#Only looking at distance to edge
prob_burn_area_DTE <- glmer(burned ~ distance_to_edge + (1|band_numb),
                 data = summer_locs_veg_data_modified,
                 na.action = "na.pass",
                 family = "binomial")
 
#Viewing summary 
summary(prob_burn_area_DTE)
```


Lets try our candidate models from our proposal.
```{r}
globalmodel <- glmer(burned ~ median_percent_grasses + I(median_percent_grasses^2) +
                               median_percent_forbs + I(median_percent_forbs^2) +
                               median_percent_shrubs + I(median_percent_shrubs^2) +
                               median_max_grasses + I(median_max_grasses^2) +
                               median_max_forbs + I(median_max_forbs^2) +
                               median_max_shrubs + I(median_max_shrubs^2) +
                               distance_to_edge + I(distance_to_edge^2) +
                               DSF + I(DSF^2) + 
                               (1|band_numb),
                    data = summer_locs_veg_data_modified,
                    na.action = "na.pass",
                    family = "binomial")

model1 <- glmer(burned ~ median_max_shrubs + (1|band_numb),
                    data = summer_locs_veg_data_modified,
                    na.action = "na.pass",
                    family = "binomial")

model2 <- glmer(burned ~ median_max_shrubs + I(median_max_shrubs^2) + (1|band_numb),
                    data = summer_locs_veg_data_modified,
                    na.action = "na.pass",
                    family = "binomial")

model3 <- glmer(burned ~ median_percent_shrubs + (1|band_numb),
                    data = summer_locs_veg_data_modified,
                    na.action = "na.pass",
                    family = "binomial")

model4 <- glmer(burned ~ median_percent_shrubs + I(median_percent_shrubs^2) + (1|band_numb),
                    data = summer_locs_veg_data_modified,
                    na.action = "na.pass",
                    family = "binomial")


model5 <-  glmer(burned ~ median_max_grasses +  median_max_forbs + (1|band_numb),
                    data = summer_locs_veg_data_modified,
                    na.action = "na.pass",
                    family = "binomial")

model6 <- glmer(burned ~ median_percent_grasses +  median_percent_forbs + (1|band_numb),
                    data = summer_locs_veg_data_modified,
                    na.action = "na.pass",
                    family = "binomial")

model7 <- glmer(burned ~ median_max_grasses +  median_max_forbs + median_max_shrubs + (1|band_numb),
                    data = summer_locs_veg_data_modified,
                    na.action = "na.pass",
                    family = "binomial")

model8 <- glmer(burned ~  median_max_grasses +  median_max_forbs + median_max_shrubs + distance_to_edge + (1|band_numb),
                    data = summer_locs_veg_data_modified,
                    na.action = "na.pass",
                    family = "binomial")

model9 <- glmer(burned ~ DSF + (1|band_numb),
                    data = summer_locs_veg_data_modified,
                    na.action = "na.pass",
                    family = "binomial")

model10 <- glmer(burned ~ DSF + I(DSF^2) + (1|band_numb),
                    data = summer_locs_veg_data_modified,
                    na.action = "na.pass",
                    family = "binomial")

model11 <-  glmer(burned ~distance_to_edge + (1|band_numb),
                    data = summer_locs_veg_data_modified,
                    na.action = "na.pass",
                    family = "binomial")

model12 <- glmer(burned ~ distance_to_edge + I(distance_to_edge^2) + (1|band_numb),
                    data = summer_locs_veg_data_modified,
                    na.action = "na.pass",
                    family = "binomial")


model13 <- glmer(burned ~ DSF + distance_to_edge + (1|band_numb),
                    data = summer_locs_veg_data_modified,
                    na.action = "na.pass",
                    family = "binomial")

model14 <- glmer(burned ~ distance_to_edge + I(distance_to_edge^2) + DSF + (1|band_numb),
                    data = summer_locs_veg_data_modified,
                    na.action = "na.pass",
                    family = "binomial")

model15 <- glmer(burned ~ DSF + I(DSF^2) + distance_to_edge + (1|band_numb),
                    data = summer_locs_veg_data_modified,
                    na.action = "na.pass",
                    family = "binomial")

model16 <- glmer(burned ~ median_percent_grasses + median_max_grasses +(1|band_numb),
                    data = summer_locs_veg_data_modified,
                    na.action = "na.pass",
                    family = "binomial")

model17<- glmer(burned ~ median_max_grasses + median_max_forbs + median_max_shrubs + distance_to_edge + DSF + (1|band_numb),
                    data = summer_locs_veg_data_modified,
                    na.action = "na.pass",
                    family = "binomial")

model18 <- glmer(burned ~ median_percent_shrubs + distance_to_edge + DSF + (1|band_numb),
                    data = summer_locs_veg_data_modified,
                    na.action = "na.pass",
                    family = "binomial")

model19 <- glmer(burned ~ median_percent_shrubs + I(median_percent_shrubs^2) + distance_to_edge + DSF + (1|band_numb),
                    data = summer_locs_veg_data_modified,
                    na.action = "na.pass",
                    family = "binomial")

model20 <- glmer(burned ~ median_max_shrubs + distance_to_edge + DSF + (1|band_numb),
                    data = summer_locs_veg_data_modified,
                    na.action = "na.pass",
                    family = "binomial")

model21 <- glmer(burned ~ median_max_shrubs + I(median_max_shrubs^2) + distance_to_edge + DSF + (1|band_numb),
                    data = summer_locs_veg_data_modified,
                    na.action = "na.pass",
                    family = "binomial")

model22 <- glmer(burned ~ median_max_grasses + median_max_forbs + distance_to_edge + DSF + (1|band_numb),
                    data = summer_locs_veg_data_modified,
                    na.action = "na.pass",
                    family = "binomial")

model23 <- glmer(burned ~ median_percent_grasses + median_percent_forbs + distance_to_edge + DSF + (1|band_numb),
                    data = summer_locs_veg_data_modified,
                    na.action = "na.pass",
                    family = "binomial")

```

Ranking candidate models by AIC
```{r}
# Create a data frame to store model names and their AIC values
model_names <- c("globalmodel", "model1", "model2", "model3", "model4",
                 "model5", "model6", "model7", "model8", "model9",
                 "model10", "model11", "model12", "model13", "model14",
                 "model15", "model16", "model17", "model18", "model19",
                 "model20", "model21", "model22", "model23")

model_list <- list(globalmodel, model1, model2, model3, model4,
                   model5, model6, model7, model8, model9,
                   model10, model11, model12, model13, model14,
                   model15, model16, model17, model18, model19,
                   model20, model21, model22, model23)

model_aic <- sapply(model_list, AIC)

# Combine the model names and AIC values into a data frame
model_comparison <- data.frame(Model = model_names, AIC = model_aic)

# Sort the models by AIC value
model_comparison_sorted <- model_comparison %>% arrange(AIC)

# View the sorted data frame
model_comparison_sorted %>% view()

# Calculate the minimum AIC value among all models
min_aic <- min(model_comparison_sorted$AIC)

# Add a column for Delta AIC
model_comparison_sorted <- model_comparison_sorted %>%
  mutate(Delta_AIC = AIC - min_aic)

# View the updated data frame
model_comparison_sorted %>% view()
```

