---
title: "Summer2023_UseAfterRXfire"
output: html_document
date: "2023-09-27"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
#### Libraries ####

library(janitor)
library(tidyverse)
library(ggthemes)
library(lubridate)
library(lme4)
library(readxl)
library(writexl)
library(sf)
library(MuMIn)
library(scales)
library(sjPlot)
library(rsq)
library(betareg)
library(broom)
library(caret)
library(janitor)
library(performance)
library(gam)
library(mgcv)
theme_set(theme_minimal())
```

Importing data. Veg metrics and percent burned already added in 2023 Burned Unburned Locs Veg .rmd
```{r}
#### Importing data ####
s23rx <- read_xlsx("/Volumes/Samsung_T5/BOBWHITE_DATA/Clean/SummerLocations/Summer2023/s23rx.xlsx") %>%
  rename(pg = perc_grasses,
         pf = perc_forbs,
         ps = perc_shrubs,
         hg = height_grasses,
         hf = height_forbs,
         hs = height_shrubs,
         pib = percent_in_burned)
```

Plotting use over time
```{r}
#### Plotting use over time ####

s23rx %>%
  dplyr::filter(date <= "2023-08-04") %>%
  ggplot(aes(dsf, pib)) +
  geom_point() +
  labs(title = "2023: Use of burned areas following prescribed fire",
       x = "DSF",
       y = "% in burned")


```

Centering and scaling veg covariates
```{r}
s23rx_sc <- s23rx %>% 
  mutate_at(vars(pg:hs), .funs = function(x){as.numeric(scale(x, center = T))}) %>%
  mutate(pib = pib + 0.00001)
```

Candidate models (betareg)
```{r}
#### Candidate models with betareg (post meeting) ####
#Global model
bmodel1 <- betareg(pib ~ pg + pf + ps + hg + hf + dsf,
                     data = s23rx_sc,
                     na.action = "na.pass")


bmodel2 <- betareg(pib ~ dsf, 
                  data = s23rx_sc,
                  na.action = "na.pass")

bmodel3<- betareg(pib ~ dsf + I(dsf^2), 
                  data = s23rx_sc,
                  na.action = "na.pass")

bmodel4 <- betareg(pib ~ pg + pf + ps,
                  data = s23rx_sc,
                  na.action = "na.pass")

bmodel5 <- betareg(pib ~ hg + hf + hs,
                  data = s23rx_sc,
                  na.action = "na.pass")


bmodel6 <- betareg(pib ~ pg + hg + pf + hg,
                  data = s23rx_sc,
                  na.action = "na.pass")

bmodel7 <- betareg(pib ~ ps,
                  data = s23rx_sc,
                  na.action = "na.pass")

bmodel8 <- betareg(pib ~ 1,
                  data = s23rx_sc,
                  na.action = "na.pass")

```

Ranking candidate models (betareg)
```{r}

#### Ranking betareg models ####
# Creating a data frame to store model names and their AIC values
model_names <- c(paste0("bmodel", 1:8))


model_list <- list(bmodel1, bmodel2, bmodel3, bmodel4,
                   bmodel5, bmodel6, bmodel7, bmodel8)


model_aic <- sapply(model_list, AIC)
model_nll <- sapply(model_list, logLik)

# Combine the model names and AIC values into a data frame
model_comparison <- data.frame(Model = model_names, AIC = model_aic, nll = - model_nll)

# Sort the models by AIC value
model_comparison_sorted <- model_comparison %>% 
  arrange(AIC)

# View the sorted data frame
model_comparison_sorted %>% 
  view()

# Calculate the minimum AIC value among all models
min_aic <- min(model_comparison_sorted$AIC)

# Add a column for Delta AIC
model_comparison_sorted <- model_comparison_sorted %>%
  mutate(Delta_AIC = AIC - min_aic)

# View the updated data frame
model_comparison_sorted %>% view()

model_comparison_sorted$rll <- exp(-0.5*model_comparison_sorted$Delta_AIC)
model_comparison_sorted$weight <- model_comparison_sorted$rll/sum(model_comparison_sorted$rll)
model_comparison_sorted$weight <- round(model_comparison_sorted$weight,3)

model_comparison_sorted %>% view
```

Candidate models (GAM)
```{r}
#### Candidate models (GAM) 2023 #### 

gmodel1 <- gam(pib ~ pg + pf + ps + hg + hf + dsf,
                     data = s23rx_sc,
                     family=betar(link = "logit"),
                     na.action = "na.pass")


gmodel2 <- gam(pib ~ dsf, 
                  data = s23rx_sc,
                  family = betar(link = "logit"),
                  na.action = "na.pass")

gmodel3 <- gam(pib ~ dsf + I(dsf^2), 
                  data = s23rx_sc,
                  family = betar(link = "logit"),
                  na.action = "na.pass")

gmodel4 <- gam(pib ~ pg + pf + ps,
                  data = s23rx_sc,
                  family = betar(link = "logit"),
                  na.action = "na.pass")

gmodel5 <- gam(pib ~ hg + hf + hs,
                  family = betar(link = "logit"),
                  data = s23rx_sc,
                  na.action = "na.pass")


gmodel6 <- gam(pib ~ pg + hg + pf + hg,
                  family = betar(link = "logit"),
                  data = s23rx_sc,
                  na.action = "na.pass")

gmodel7 <- gam(pib ~ ps,
                   family = betar(link = "logit"),
                  data = s23rx_sc,
                  na.action = "na.pass")

gmodel8 <- gam(pib ~ 1,
                  family = betar(link = "logit"),
                  data = s23rx_sc,
                  na.action = "na.pass")
```


Ranking candidate models (gam)
```{r}

#### Ranking ga models ####
# Creating a data frame to store model names and their AIC values
model_names <- c(paste0("gmodel", 1:8))


model_list <- list(gmodel1, gmodel2, gmodel3, gmodel4,
                   gmodel5, gmodel6, gmodel7, gmodel8)


model_aic <- sapply(model_list, AIC)
model_nll <- sapply(model_list, logLik)

# Combine the model names and AIC values into a data frame
model_comparison <- data.frame(Model = model_names, AIC = model_aic, nll = - model_nll)

# Sort the models by AIC value
model_comparison_sorted <- model_comparison %>% 
  arrange(AIC)

# View the sorted data frame
model_comparison_sorted %>% 
  view()

# Calculate the minimum AIC value among all models
min_aic <- min(model_comparison_sorted$AIC)

# Add a column for Delta AIC
model_comparison_sorted <- model_comparison_sorted %>%
  mutate(Delta_AIC = AIC - min_aic)

# View the updated data frame
model_comparison_sorted %>% view()

model_comparison_sorted$rll <- exp(-0.5*model_comparison_sorted$Delta_AIC)
model_comparison_sorted$weight <- model_comparison_sorted$rll/sum(model_comparison_sorted$rll)
model_comparison_sorted$weight <- round(model_comparison_sorted$weight,3)

model_comparison_sorted %>% view
```

Lets try combining 2023/23 data
```{r}
s22_select <- s22rx_sc %>%
  select(DSF, pg:hs, pib)

s22_select <- s22_select %>%
  rename(dsf = DSF)

s23_select <- s23rx_sc %>%
  select(dsf, pg:hs, pib)

s22_23_rx <- rbind(
  s22_select, s23_select
)
```

Candidate models (GAM) 2022/23
```{r}
#### Candidate models (GAM)  FOR BOTH YEARS #### 

gmodel1 <- gam(pib ~ pg + pf + ps + hg + hf + dsf,
                     data = s22_23_rx,
                     family=betar(link = "logit"),
                     na.action = "na.pass")


gmodel2 <- gam(pib ~ dsf, 
                  data = s22_23_rx,
                  family = betar(link = "logit"),
                  na.action = "na.pass")

gmodel3 <- gam(pib ~ dsf + I(dsf^2), 
                  data = s22_23_rx,
                  family = betar(link = "logit"),
                  na.action = "na.pass")

gmodel4 <- gam(pib ~ pg + pf + ps,
                  data = s22_23_rx,
                  family = betar(link = "logit"),
                  na.action = "na.pass")

gmodel5 <- gam(pib ~ hg + hf + hs,
                  family = betar(link = "logit"),
                  data = s22_23_rx,
                  na.action = "na.pass")


gmodel6 <- gam(pib ~ pg + hg + pf + hg,
                  family = betar(link = "logit"),
                  data = s22_23_rx,
                  na.action = "na.pass")

gmodel7 <- gam(pib ~ ps,
                   family = betar(link = "logit"),
                  data = s22_23_rx,
                  na.action = "na.pass")

gmodel8 <- gam(pib ~ 1,
                  family = betar(link = "logit"),
                  data = s22_23_rx,
                  na.action = "na.pass")
```


Ranking candidate models (gam)
```{r}

#### Ranking gam models 2022/23 ####
# Creating a data frame to store model names and their AIC values
model_names <- c(paste0("gmodel", 1:8))


model_list <- list(gmodel1, gmodel2, gmodel3, gmodel4,
                   gmodel5, gmodel6, gmodel7, gmodel8)


model_aic <- sapply(model_list, AIC)
model_nll <- sapply(model_list, logLik)

# Combine the model names and AIC values into a data frame
model_comparison <- data.frame(Model = model_names, AIC = model_aic, nll = - model_nll)

# Sort the models by AIC value
model_comparison_sorted <- model_comparison %>% 
  arrange(AIC)

# View the sorted data frame
model_comparison_sorted %>% 
  view()

# Calculate the minimum AIC value among all models
min_aic <- min(model_comparison_sorted$AIC)

# Add a column for Delta AIC
model_comparison_sorted <- model_comparison_sorted %>%
  mutate(Delta_AIC = AIC - min_aic)

# View the updated data frame
model_comparison_sorted %>% view()

model_comparison_sorted$rll <- exp(-0.5*model_comparison_sorted$Delta_AIC)
model_comparison_sorted$weight <- model_comparison_sorted$rll/sum(model_comparison_sorted$rll)
model_comparison_sorted$weight <- round(model_comparison_sorted$weight,3)

model_comparison_sorted %>% view
```

Single variable candidate models (GAM)
```{r}
#### Candidate models (GAM) 2023 #### 

gmodel1 <- gam(pib ~ pg,
                     data = s23rx_sc,
                     family=betar(link = "logit"),
                     na.action = "na.pass")


gmodel2 <- gam(pib ~ dsf, 
                  data = s23rx_sc,
                  family = betar(link = "logit"),
                  na.action = "na.pass")

gmodel3 <- gam(pib ~ pf, 
                  data = s23rx_sc,
                  family = betar(link = "logit"),
                  na.action = "na.pass")

gmodel4 <- gam(pib ~ ps,
                  data = s23rx_sc,
                  family = betar(link = "logit"),
                  na.action = "na.pass")

gmodel5 <- gam(pib ~ hg,
                  family = betar(link = "logit"),
                  data = s23rx_sc,
                  na.action = "na.pass")


gmodel6 <- gam(pib ~ hf,
                  family = betar(link = "logit"),
                  data = s23rx_sc,
                  na.action = "na.pass")

gmodel7 <- gam(pib ~ hs,
                   family = betar(link = "logit"),
                  data = s23rx_sc,
                  na.action = "na.pass")

gmodel8 <- gam(pib ~ 1,
                  family = betar(link = "logit"),
                  data = s23rx_sc,
                  na.action = "na.pass")
```

Ranking candidate models (gam)
```{r}

#### Ranking gam models 2022/23 ####
# Creating a data frame to store model names and their AIC values
model_names <- c(paste0("gmodel", 1:8))


model_list <- list(gmodel1, gmodel2, gmodel3, gmodel4,
                   gmodel5, gmodel6, gmodel7, gmodel8)


model_aic <- sapply(model_list, AIC)
model_nll <- sapply(model_list, logLik)

# Combine the model names and AIC values into a data frame
model_comparison <- data.frame(Model = model_names, AIC = model_aic, nll = - model_nll)

# Sort the models by AIC value
model_comparison_sorted <- model_comparison %>% 
  arrange(AIC)

# View the sorted data frame
model_comparison_sorted %>% 
  view()

# Calculate the minimum AIC value among all models
min_aic <- min(model_comparison_sorted$AIC)

# Add a column for Delta AIC
model_comparison_sorted <- model_comparison_sorted %>%
  mutate(Delta_AIC = AIC - min_aic)

# View the updated data frame
model_comparison_sorted %>% view()

model_comparison_sorted$rll <- exp(-0.5*model_comparison_sorted$Delta_AIC)
model_comparison_sorted$weight <- model_comparison_sorted$rll/sum(model_comparison_sorted$rll)
model_comparison_sorted$weight <- round(model_comparison_sorted$weight,3)

model_comparison_sorted %>% view
```