---
title: "Summer2022UseAfterRXfire"
output: html_document
date: "2023-09-21"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Loading libraries
```{r}
#### Libraries ####

library(janitor)
library(tidyverse)
library(ggthemes)
library(lubridate)
library(lme4)
library(readxl)
library(writexl)
library(sf)
library(MuMIn)
library(scales)
library(sjPlot)
library(rsq)
library(betareg)
library(broom)
library(caret)
library(janitor)
library(performance)
library(gam)
library(mgcv)
theme_set(theme_minimal())
```

Importing locations
```{r}
#### Importing locations ####

#Importing summer locations, already joined with veg data and birds online.
summer2022_locs_veg_data <- read_xlsx("/Volumes/Samsung_T5/BOBWHITE_DATA/Clean/SummerLocations/Summer2022/summer_2022_points_veg_CorrectBirdsOnline.xlsx") %>%
  mutate(date = ymd(date)) %>%
  dplyr::filter(status != "D") 

#Adding in age/sex of each bird. May be useful
#The fate spreadsheet has age and sex info for each bird.
#Importing "fate" spreadsheet
fate <- rio::import("/Volumes/Samsung_T5/BOBWHITE_DATA/Clean/Fate/fate_clean_2021_22.xlsx", setclass = "tibble") %>%
  clean_names() %>%
  rename(band_numb = band_number) 

#Now we bind age and sex into locations, based on band_numb
summer2022_locs_veg_data <- summer2022_locs_veg_data %>%
  left_join(fate %>% 
              dplyr::select(band_numb, age, sex), 
            by = "band_numb")

veg_data_2022 <- read_xlsx("/Volumes/Samsung_T5/BOBWHITE_DATA/Clean/VegSurveys/VegSurveys2022.xlsx") %>%
  clean_names()

```

Determining proportion of birds in burn areas each day.
```{r}
#### Determining proportion of birds located in burned areas daily ####
# Counting unique birds signaled within a burned area each day
birds_in_burned_areas <- summer2022_locs_veg_data %>%
  dplyr::filter(burned == 1) %>%
  group_by(date) %>%
  summarize(birds_in_burned_areas = n_distinct(band_numb))


# Joining into summer locs data
summer2022_locs_veg_data <- summer2022_locs_veg_data %>%
  left_join(birds_in_burned_areas, by = "date")

# Handling NA values for birds_in_burned_areas
summer2022_locs_veg_data$birds_in_burned_areas[is.na(summer2022_locs_veg_data$birds_in_burned_areas)] <- 0

# Calculate percentage of birds in burned areas
s22rx <- summer2022_locs_veg_data %>%
  mutate(pib = (birds_in_burned_areas / birds_online))

# Handling NA values for pib
s22rx$pib[is.na(s22rx$pib)] <- 0


ave_burned_per_week <- s22rx %>%
  group_by(week) %>%
  summarize(mean_per_burned = mean(pib))
```


Lets add a distance to edge column for our points 
```{r}
#### Importing map of burned areas, calculating distance ####

#Importing map of burned areas, summer 2022.
#Reading burn compartments in again, but as an sf object
burn_compartments_2022 <- st_read("/Volumes/Samsung_T5/BOBWHITE_DATA/Maps/2022_burn_compartments/Polygons/NoAG_NoHW/2022_burn_units_NoAG_NoHW.shp")

# Converting burn_compartments_2022 polygons to line geometries that represent their boundaries. This is necessary to calc min distances later.
burn_compartment_boundaries <- st_boundary(burn_compartments_2022)

# Converting tibble of points to an sf object
summer2022_locs_sf <- st_as_sf(s22rx, coords = c("easting", "northing"), crs = st_crs(burn_compartments_2022))

# Calculating the distance from each point to the nearest line boundary of the burn compartments, storing it in matrix
distance_matrix <- st_distance(summer2022_locs_sf, burn_compartment_boundaries)

# Applying the function min() to each row (1 does this -- we'd use 2 for columns) in the distance matrix
min_distances <- apply(distance_matrix, 1, min)

# Add the minimum distance to the original tibble as a new column
s22rx$distance_to_edge <- min_distances


#This looks right, and similar to what we got in other analysis.
s22rx %>%
  dplyr::filter(burned == 1) %>%
  ggplot(aes(DSF, distance_to_edge)) +
  geom_point() +
  labs(x = "Days since fire",
       y = "Distance to perimeter of burn unit (m)") +
  theme_bw() +
  theme(legend.position = "none",
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line("black"),
        axis.ticks = element_line("black"),
        axis.text.x = element_text(colour = "black", size = 12),
        axis.title.y = element_text(colour = "black", size = 12),
        axis.title.x = element_text(colour = "black", size = 12),
        legend.title = element_text(colour = "black", size = 12),
        legend.text = element_text(colour = "black", size = 12),
        axis.text.y = element_text(colour = "black", size = 12),
        plot.margin = margin(t = 5, r = 10, b = 5, l = 5, unit = "pt"))

dist_model <- lmer(distance_to_edge ~ DSF + (1 | band_numb), 
                    data = s22rx %>% dplyr::filter(burned == 1))
summary(dist_model)

s22rx %>%
  dplyr::filter(burned == 1) %>%
  summarize(mean_distance = mean(distance_to_edge))


s22rx %>%
  dplyr::filter(burned == 1) %>%
  ggplot(aes(distance_to_edge)) +
  geom_histogram(binwidth = 10, fill = "midnight blue", alpha = 0.8) + 
  geom_vline(xintercept = 28, linetype = "dashed", color = "gray") +
  labs(x = "Distance to edge",
       y = "",
       title = "Distances to edge following prescribed fire",
       subtitle = "Only looking at birds within burn compartments")

  

s22rx %>%
  dplyr::filter(burned == 0) %>%
  ggplot(aes(date, distance_to_edge)) +
  geom_point() +
  geom_smooth() +
  labs(x = "DSF",
       y = "Distance to edge",
       title = "Distance to edge of burned compartments following prescribed fire",
       subtitle = "Only looking at birds located in unburned units")

s22rx %>%
  dplyr::filter(burned == 0) %>%
  summarize(mean_distance = mean(distance_to_edge))


s22rx %>%
  dplyr::filter(burned == 0) %>%
  ggplot(aes(distance_to_edge)) +
  geom_histogram(binwidth = 20, fill = "midnight blue", alpha = 0.8) + 
  geom_vline(xintercept = 95.8, linetype = "dashed", color = "gray") +
  labs(x = "Distance to edge",
       y = "",
       title = "Distances to edge following prescribed fire",
       subtitle = "Only looking at birds in unburned compartments")

s22rx %>%
  mutate(burned = factor(burned, labels = c("Unburned", "Burned"))) %>%
  ggplot(aes(x = burned, y = distance_to_edge)) +
  geom_boxplot(fill = "midnight blue", alpha = 0.5) +
  labs(x = "Burn Status",
       y = "Distance to Edge",
       title = "Distance to edge of burn compartments following prescribed fire")

s22rx %>%
  mutate(burned = factor(burned, labels = c("Unburned", "Burned"))) %>%
  ggplot(aes(x = burned, y = distance_to_edge)) +
  geom_violin(fill = "midnight blue", alpha = 0.5) +
  labs(x = "Burn Status",
       y = "Distance to Edge",
       title = "Distance to edge following prescribed fire")

s22dist <- s22rx %>%
  select(burned, distance_to_edge)

```


Making some plots of use of burned areas over time.
```{r}
#### Plotting use over time ####

#Average proportion of birds in burn areas each day, for each week.
ave_burned_per_week %>%
  mutate(week = (week - 15)) %>%
  dplyr::filter(week <=15) %>%
  ggplot(aes(week, mean_per_burned)) +
  geom_line() +
  labs(x = "Weeks following prescribed fire",
       y = "Mean percentage of birds online that were located in burned areas",
       title = "Average percentage of birds online that were located in burned areas each week")

#proportion of birds in burn areas each day.
s22rx_uniq %>%
  ggplot(aes(DSF, pib)) +
  geom_point(size = 2) +
  labs(x = "Days since fire",
       y = "Proportion of bobwhite in burned areas") +
  theme_bw() +
  theme(legend.position = "none",
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line("black"),
        axis.ticks = element_line("black"),
        axis.text.x = element_text(colour = "black", size = 12),
        axis.title.y = element_text(colour = "black", size = 12),
        axis.title.x = element_text(colour = "black", size = 12),
        legend.title = element_text(colour = "black", size = 12),
        legend.text = element_text(colour = "black", size = 12),
        axis.text.y = element_text(colour = "black", size = 12),
        plot.margin = margin(t = 5, r = 10, b = 5, l = 5, unit = "pt")) 

```

Centering and scaling vegetation covariates, removing duplicate rows.
```{r}
#### Centering and scaling veg covariates. Removing duplicate rows ####

s22rx_sc <- s22rx %>% 
  mutate_at(vars(pg:hs), .funs = function(x){as.numeric(scale(x, center = T))}) %>%
  mutate(pib = pib + 0.00001)

s22rx_sc <- s22rx_sc %>%
  select(pib, DSF, pg:hs)

s22rx_uniq <- s22rx_sc %>% 
  group_by(DSF) %>%
  slice_head(n = 1) %>%
  ungroup()

View(s22rx_uniq)
```

Checking for correlations 
```{r}
#### Checking for correlations ####
veg_only <- read_xlsx("/Volumes/Samsung_T5/BOBWHITE_DATA/Clean/VegSurveys/VegSurveys2022.xlsx") %>% 
  clean_names %>%
  select(percent_grasses:max_shrubs_cm)

#Before centering and scaling
cor(veg_only)

#After centering and scaling
csveg <- s22rx_sc %>%
  select(pg:hs)

cor(csveg)
```

Adding in temp data
```{r}
temp22 <- read_csv("/Volumes/Samsung_T5/BOBWHITE_DATA/Clean/Other/IchTempData.csv") %>%
  clean_names() %>%
  mutate(date = as.Date(date, tryFormats = "%m-%d-%Y")) 

temp22 <- temp22[temp22$date %in% seq(as.Date("2022-04-11"), by="day",length.out=116),]

start_date22 <- ymd("2022-04-11")

temp22$DSF <- as.numeric(temp22$date - start_date22)


s22_rx_uniq_temp <- s22rx_uniq %>%
  left_join(temp22, by = "DSF") %>%
  select(pib:date, air_temp_max_c)

s22_rx_uniq_temp$air_temp_max_c <- ifelse(is.na(s22_rx_uniq_temp$air_temp_max_c), 27.2, s22_rx_uniq_temp$air_temp_max_c)

s22_rx_uniq_temp$year <- 2022

s22_rx_uniq_temp_mod <- s22_rx_uniq_temp %>%
  mutate_at(vars(air_temp_max_c), .funs = function(x){as.numeric(scale(x, center = T))})

```



Univarariate candidate models (GAM)
```{r}
#### Candidate models (GAM) 2022 #### 

gmodel1 <- gam(pib ~ pg,
                     data = s22rx_uniq,
                     family=betar(link = "logit"),
                     na.action = "na.pass")


gmodel2 <- gam(pib ~ DSF, 
                  data = s22rx_uniq,
                  family = betar(link = "logit"),
                  na.action = "na.pass")

gmodel3 <- gam(pib ~ pf, 
                  data = s22rx_uniq,
                  family = betar(link = "logit"),
                  na.action = "na.pass")

gmodel4 <- gam(pib ~ ps,
                  data = s22rx_uniq,
                  family = betar(link = "logit"),
                  na.action = "na.pass")

gmodel5 <- gam(pib ~ hg,
                  family = betar(link = "logit"),
                  data = s22rx_uniq,
                  na.action = "na.pass")


gmodel6 <- gam(pib ~ hf,
                  family = betar(link = "logit"),
                  data = s22rx_uniq,
                  na.action = "na.pass")

gmodel7 <- gam(pib ~ hs,
                   family = betar(link = "logit"),
                  data = s22rx_uniq,
                  na.action = "na.pass")

gmodel8 <- gam(pib ~ 1,
                  family = betar(link = "logit"),
                  data = s22rx_uniq,
                  na.action = "na.pass")

#gmodel9 <- gam(pib ~  DSF + air_temp_max_c,
                #  family = betar(link = "logit"),
                #  data = s22_rx_uniq_temp_mod,
               #  na.action = "na.pass")

#gmodel10 <- gam(pib ~ air_temp_max_c,
                 # family = betar(link = "logit"),
                 # data = s22_rx_uniq_temp_mod,
                 # na.action = "na.pass")

#summary(gmodel9)

```

Ranking candidate models (gam)
```{r}

#### Ranking gam models 2022 ####
# Creating a data frame to store model names and their AIC values
model_names <- c(paste0("gmodel", 1:8))


model_list <- list(gmodel1, gmodel2, gmodel3, gmodel4,
                   gmodel5, gmodel6, gmodel7, gmodel8)


model_aic <- sapply(model_list, AIC)
model_nll <- sapply(model_list, logLik)

# Combine the model names and AIC values into a data frame
model_comparison <- data.frame(Model = model_names, AIC = model_aic, nll = - model_nll)

# Sort the models by AIC value
model_comparison_sorted <- model_comparison %>% 
  arrange(AIC)

# View the sorted data frame
model_comparison_sorted %>% 
  view()

# Calculate the minimum AIC value among all models
min_aic <- min(model_comparison_sorted$AIC)

# Add a column for Delta AIC
model_comparison_sorted <- model_comparison_sorted %>%
  mutate(Delta_AIC = AIC - min_aic)

# View the updated data frame
model_comparison_sorted %>% view()

model_comparison_sorted$rll <- exp(-0.5*model_comparison_sorted$Delta_AIC)
model_comparison_sorted$weight <- model_comparison_sorted$rll/sum(model_comparison_sorted$rll)
model_comparison_sorted$weight <- round(model_comparison_sorted$weight,3)

model_comparison_sorted %>% view

#write_xlsx(model_comparison_sorted, "/Users/jeffgrayum/Downloads/model_comp22FIXED.xlsx")
```

```{r}
##### Now for DSF predict ####
DSF_values <- seq(min(s22rx_uniq$DSF, na.rm = TRUE), 
                 max(s22rx_uniq$DSF, na.rm = TRUE), 
                 length.out = 1000)

DSF_df <- data.frame(
  hs = rep(mean(s22rx_uniq$hs), length(DSF_values)),
  hg = rep(mean(s22rx_uniq$hg), length(DSF_values)),
  hf = rep(mean(s22rx_uniq$hf), length(DSF_values)),
  pg = rep(mean(s22rx_uniq$pg), length(DSF_values)),
  pf = rep(mean(s22rx_uniq$pf), length(DSF_values)),
  ps = rep(mean(s22rx_uniq$ps), length(DSF_values)),
  DSF =  DSF_values
)

DSF_pred <- data.frame(predict(gmodel2, newdata = DSF_df, type = "response", se.fit = TRUE))


DSF_pred$lowCI = (DSF_pred$fit - (1.96 * DSF_pred$se.fit))
DSF_pred$upperCI = (DSF_pred$fit + (1.96 * DSF_pred$se.fit))


DSF_pred %>%
  ggplot(aes(DSF_values, y = fit)) +
  geom_line() +
  geom_ribbon(aes(ymin = lowCI, ymax = upperCI), fill = "red", alpha = 0.2) +
  labs(x = "Days since fire",
       y = "Proportion of bobwhite predicted in burned areas") +
  theme_bw() +
 theme(legend.position = "none",
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line("black"),
        axis.ticks = element_line("black"),
        axis.text.x = element_text(colour = "black", size = 12),
        axis.title.y = element_text(colour = "black", size = 12),
        axis.title.x = element_text(colour = "black", size = 12),
        legend.title = element_text(colour = "black", size = 12),
        legend.text = element_text(colour = "black", size = 12),
        axis.text.y = element_text(colour = "black", size = 12),
        plot.margin = margin(t = 5, r = 10, b = 5, l = 5, unit = "pt"))
```

```{r}
#### Top model (9) predictions
#Temp first
temp_values <- seq(min(s22_rx_uniq_temp_mod$air_temp_max_c, na.rm = TRUE), 
                 max(s22_rx_uniq_temp_mod$air_temp_max_c, na.rm = TRUE), 
                 length.out = 1000)

temp_df <- data.frame(
 air_temp_max_c = temp_values,
 DSF =  rep(mean(s22_rx_uniq_temp_mod$DSF), length(temp_values))
)

temp_pred <- data.frame(predict(gmodel9, newdata = temp_df, type = "response", se.fit = TRUE))


temp_pred$lowCI = (temp_pred$fit - (1.96 * temp_pred$se.fit))
temp_pred$upperCI = (temp_pred$fit + (1.96 * temp_pred$se.fit))


temp_pred %>%
  ggplot(aes(temp_values, y = fit)) +
  geom_line() +
  geom_ribbon(aes(ymin = lowCI, ymax = upperCI), fill = "red", alpha = 0.2) +
  labs(x = "Max temp celcius (centered and scaled)",
       y = "% Bobwhite predicted in burned units",
       title = "Use of burned units after prescribed fire predicted by temperature") +
  scale_y_continuous(labels = scales::percent) +
  theme_bw() 

```

Does dsf influence distance to edge?
```{r}
#### Distance to edge model #### 

dist_model22 <- lmer(distance_to_edge ~ DSF + (1 | band_numb),
                   data = s22rx)

summary(dist_model22)

```

```{r}
#### Plotting veg data vs percent in burned ####
s22rx_piv <- s22rx %>%
  select(pg:hs, pib) %>%
  pivot_longer(pg:hs, names_to = "Measurement", values_to = "Value")


s22rx_piv %>% 
  dplyr::filter(Measurement %in% c("pg", "pf", "ps")) %>%
  mutate(Measurement = factor(Measurement, levels = c("pg", "pf", "ps"))) %>%
  ggplot(aes(Value, pib)) +
  geom_point(color = "Midnight Blue") +
  facet_wrap(~ Measurement, scales = "free") +
  labs(x = "% cover",
       y = "% of birds in burned area",
       title = "Percent cover of grasses, forbs, and shrubs vs percent of birds in burned areas") +
  theme_bw()

s22rx_piv %>% 
  dplyr::filter(Measurement %in% c("hg", "hf", "hs")) %>%
  mutate(Measurement = factor(Measurement, levels = c("hg", "hf", "hs"))) %>%
  ggplot(aes(Value, pib)) +
  geom_point(color = "Midnight Blue") +
  facet_wrap(~ Measurement, scales = "free") +
  labs(x = "Height (cm)",
       y = "% of birds in burned area",
       title = "Height of grasses, forbs and shrubs vs percent of birds in burned areas") +
  theme_bw()


s22rx %>%
  ggplot(aes(pg, pib)) +
  geom_point(color = "Midnight Blue") +
  labs(x = "% cover of grasses",
       y = "Perenct of birds in burned area",
       title = "Percent cover grasses vs Percent of birds in burned areas") +
  theme_bw()

s22rx %>%
  ggplot(aes(pf, pib)) +
  geom_point(color = "Midnight Blue") +
  labs(x = "% cover of forbs",
       y = "Perenct of birds in burned area",
       title = "Percent cover forbs vs Percent of birds in burned areas") +
  theme_bw()


s22rx %>%
  ggplot(aes(ps, pib)) +
  geom_point(color = "Midnight Blue") +
  labs(x = "% cover of shrubs",
       y = "Perenct of birds in burned area",
       title = "Percent cover shrubs vs Percent of birds in burned areas")

s22rx %>%
  ggplot(aes(hg, pib)) +
  geom_point(color = "Midnight Blue") +
  labs(x = "Height of grasses (cm)",
       y = "Perenct of birds in burned area",
       title = "Height of grasses vs Percent of birds in burned areas")

s22rx %>%
  ggplot(aes(hf, pib)) +
  geom_point(color = "Midnight Blue") +
  labs(x = "Height of forbs (cm)",
       y = "Perenct of birds in burned area",
       title = "Height of forbs vs Percent of birds in burned areas")

s22rx %>%
  ggplot(aes(hs, pib)) +
  geom_point(color = "Midnight Blue") +
  labs(x = "Height of shrubs (cm)",
       y = "Perenct of birds in burned area",
       title = "Height of shrubs vs Percent of birds in burned areas")
  
```

Making predict plots with top model
```{r}
#### Making predict plots ####

#First, we will calculate the mean values for hg, hf, and hs. Two of these will be held constant when making predict plots for other covariate.
mean_hg <- mean(s22rx_sc$hg)
mean_hf <- mean(s22rx_sc$hf)
mean_hs <- mean(s22rx_sc$hs)

#First we will do height of grasses. We need to create a new data frame with sequence of grass height(between min a max), and the mean values of hf and hs.

# Generate a sequence of hg values, between min and max.
hg_values <- seq(min(s22rx_sc$hg, na.rm = TRUE), 
                 max(s22rx_sc$hg, na.rm = TRUE), 
                 length.out = 1000) 


# we must create a data frame with the hg values, and the mean hf and hs values
hgdf <- data.frame(
  hg = hg_values, 
  hf = rep(mean_hf, length(hg_values)), 
  hs = rep(mean_hs, length(hg_values))
)

# Generating predictions and variance of predictions.
hg_pred <- data.frame(predict(model5, newdata = hgdf, type = "response", se.fit = TRUE))

hg_pred$hg_values <- hg_values

#head(hg_pred)

hg_pred$lowCI = (hg_pred$fit - (1.96 * hg_pred$se.fit))
hg_pred$upperCI = (hg_pred$fit + (1.96 * hg_pred$se.fit))

hg_pred %>%
  ggplot(aes(hg_values, fit)) +
  geom_line() +
  geom_ribbon(aes(ymin = lowCI, ymax = upperCI), fill = "blue", alpha = 0.1) +
  labs(x = "Height of grasses (centered and scaled)",
       y = "Predicted % NOBO in burned areas",
       title = "Predicted % NOBO in burned areas vs height of grasses") +
  scale_y_continuous(labels = percent)

#Now, we will do hf.
#Generate a sequence of hf values, between min and max.
hf_values <- seq(min(s22rx_sc$hf, na.rm = TRUE), 
                 max(s22rx_sc$hf, na.rm = TRUE), 
                 length.out = 300) 


#We must create a data frame with the hg values, and the mean hf and hs values
hfdf <- data.frame(
  hg = rep(mean_hg, length(hf_values)), 
  hf = hf_values, 
  hs = rep(mean_hs, length(hf_values))
)

# Generate predictions
hf_pred <- data.frame(predict(model5, newdata = hfdf, type = "response", se.fit = TRUE))

hf_pred$hf_values <- hf_values

hf_pred$lowCI = (hf_pred$fit - (1.96 * hf_pred$se.fit))
hf_pred$upperCI = (hf_pred$fit + (1.96 * hf_pred$se.fit))

# Plotting
hf_pred%>% 
  ggplot(aes(hf_values, fit)) +
  geom_line() +
  geom_ribbon(aes(ymin = lowCI, ymax = upperCI), fill = "blue", alpha = 0.1) +
  labs(x = "Height of forbs (centered and scaled)",
       y = "Predicted % NOBO in burned areas",
       title = "Predicted % NOBO in burned areas vs height of forbs") +
  scale_y_continuous(labels = percent)



#Now, we will do hs.
# Generate a sequence of hf values, between min and max.
hs_values <- seq(min(s22rx_sc$hs, na.rm = TRUE), 
                 max(s22rx_sc$hs, na.rm = TRUE), 
                 length.out = 1000) 


# we must create a data frame with the hg values, and the mean hf and hs values
hsdf <- data.frame(
  hg = rep(mean_hg, length(hs_values)), 
  hf = rep(mean_hf, length(hs_values)), 
  hs = hs_values)


# Generate predictions
hs_pred <- data.frame(predict(model5, newdata = hsdf, type = "response", se.fit = TRUE))


#Calculating upper and lower confidence intervals
hs_pred$lowCI = (hs_pred$fit - (1.96 * hs_pred$se.fit))
hs_pred$upperCI = (hs_pred$fit + (1.96 * hs_pred$se.fit))


# Plotting
hs_pred%>% 
  ggplot(aes(hs_values, fit)) +
  geom_line() +
  geom_ribbon(aes(ymin = lowCI, ymax = upperCI), fill = "blue", alpha = 0.1) +
  labs(x = "Height of shrubs (centered and scaled)",
       y = "Predicted % NOBO in burned areas",
       title = "Predicted % NOBO in burned areas vs height of shrubs") +
  scale_y_continuous(labels = percent)



#Now, we will do hs.
# Generate a sequence of hf values, between min and max.
hs_values <- seq(min(s22rx_sc$hs, na.rm = TRUE), 
                 max(s22rx_sc$hs, na.rm = TRUE), 
                 length.out = 300) 


# we must create a data frame with the hg values, and the mean hf and hs values
hsdf <- data.frame(
  hg = rep(mean_hg, length(hs_values)), 
  hf = rep(mean_hf, length(hs_values)), 
  hs = hs_values)


# Generate predictions
hs_mean <- predict(model5, newdata = hsdf, type = "response")
hs_var <- predict(model5, newdata = hsdf, type = "variance")

head(hs_mean)

#Creating new dataframe with hs mean predictions, and variance
hs_results_df <- data.frame(
  hs_mean = hs_mean,
  hs_var = hs_var
)

#Calculating upper and lower confidence intervals
hs_results_df$lowCI = (hs_results_df$hs_mean - (1.96 * sqrt(hs_results_df$hs_var)))
hs_results_df$upperCI = (hs_results_df$hs_mean + (1.96 * sqrt(hs_results_df$hs_var)))

head(hs_results_df)

# Plotting
hs_results_df %>% 
  ggplot(aes(hs_values, hs_mean)) +
  geom_line() +
  geom_ribbon(aes(ymin = lowCI, ymax = upperCI), fill = "blue", alpha = 0.1) +
  labs(x = "Height of shrubs (centered and scaled)",
       y = "Predicted % NOBO in burned areas",
       title = "Predicted % NOBO in burned areas vs height of shrubs") +
  scale_y_continuous(labels = percent)


```

Adding in temp data
```{r}
temp22 <- read_csv("/Volumes/Samsung_T5/BOBWHITE_DATA/Clean/Other/IchTempData.csv") %>%
  clean_names() %>%
  mutate(date = as.Date(date, tryFormats = "%m-%d-%Y")) 

temp22f <- temp22[temp22$date %in% seq(as.Date("2022-04-11"), by="day",length.out=116),]

start_date22 <- ymd("2022-04-11")

temp22f$DSF <- as.numeric(temp22f$date - start_date22)

s22rx <- s22rx %>%
  left_join(temp22f, join_by = "DSF")

s22temp <- s22rx %>%
  select(DSF, air_temp_max_c)

s22temp$year <- 2022

s22temp$air_temp_max_c <- ifelse(is.na(s22temp$air_temp_max_c), 27.2, s22temp$air_temp_max_c)
```

```{r}
#### Distance to edge model #### 

#Filter for only birds found in burned areas
s22rx_bu <- s22rx %>%
  dplyr::filter(burned == 1)

dist_model22 <- lmer(distance_to_edge ~ DSF + (1 | band_numb),
                   data = s22rx_bu)

summary(dist_model22)

      
```


```{r}
s22rx %>%
  dplyr::filter(date <= "2022-08-04",
                ) %>%
  ggplot(aes(x = DSF)) +
  geom_line(aes(y = (air_temp_max_c, color = "Max Temp (C)")) +  # Divide by 2 to align with pib
  geom_point(aes(y = pib * 100, color = "% Bobwhite in Burned Area")) +
  geom_hline(yintercept = 85 / 2, linetype = "dashed", color = "blue") +  # Divide by 2 to align with pib
  scale_y_continuous(
    name = "% bobwhite in burned area",
    sec.axis = sec_axis(~ . * 1, name = "Max daily temp (C)")  # Multiply by 2 to revert the division
  ) +
  scale_color_manual(
    values = c("Max Temp (C)" = "red", "% Bobwhite in Burned Area" = "midnightblue"),
    name = ""
  ) +
  labs(x = "Days since fire",
       title = "2022: % NOBO in burned area and daily high temperature",
       subtitle = "Dashed blue line at 85* F") +
  theme_bw()

# Dummy data frame to create legend item for the blue dashed line
# Assuming s22rx exists and has the relevant columns
# Create dashed line data
# Assuming s22rx exists and has the relevant columns
# Dynamic label name
degree_label <- paste0("Max ambient temp (", "\u00B0", "F)")

# Color mapping
color_map <- setNames(c("red", "black", "blue"), c(degree_label, "% Bobwhite in burned units", "Bobwhite thermal stress threshold"))

# Assuming s22rx exists and has the relevant columns
# Create dashed line data
dashed_line_data <- data.frame(DSF = c(min(s22rx$DSF), max(s22rx$DSF)),
                               y = c(86.5 / 2, 86.5 / 2))

# Plot
s22rx %>%
  dplyr::filter(date <= "2022-08-04") %>%
  ggplot(aes(x = DSF)) +
  geom_line(aes(y = (air_temp_max_c * 1.8 + 32) / 2, color = degree_label)) +
  geom_point(aes(y = pib * 100, color = "% Bobwhite in burned units")) +
  geom_line(data = dashed_line_data, aes(y = y, color = "Bobwhite thermal stress threshold"), linetype = "dashed") +
  scale_y_continuous(
    name = "% Bobwhite in burned units",
    sec.axis = sec_axis(~ . * 2, name = degree_label)
  ) +
  scale_color_manual(
    values = color_map,
    name = ""
  ) +
  labs(x = "Days since fire",
       y = "% sampled bobwhite in burned units",
       title = paste0("Use of burned units after prescribed fire\nand max ambient temp (2022) ")) +
  theme_bw() +
  scale_x_continuous(breaks = c(0, 30, 60, 90, 120, 150))

```

```{r}

#### Making a temp data frame ####
s22_rx_uniq_temp <- s22_rx_uniq %>%
  left_join(s22temp, by = join_by(DSF), relationship = "many-to-many")

s22_rx_uniq_temp_mod <- s22_rx_uniq_temp %>%
  mutate_at(vars(air_temp_max_c), .funs = function(x){as.numeric(scale(x, center = T))})

gmodel9 <- gam(pib ~  DSF + air_temp_max_c,
                  family = betar(link = "logit"),
                  data = s22_rx_uniq_temp_mod,
                  na.action = "na.pass")

gmodel10 <- gam(pib ~  air_temp_max_c,
                  family = betar(link = "logit"),
                  data = s22_rx_uniq_temp_mod,
                  na.action = "na.pass")

summary(gmodel9)
```





