---
title: "Summer2023_UseAfterRXfire"
output: html_document
date: "2023-09-27"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
#### Libraries ####

library(janitor)
library(tidyverse)
library(ggthemes)
library(lubridate)
library(lme4)
library(readxl)
library(writexl)
library(sf)
library(MuMIn)
library(scales)
library(sjPlot)
library(rsq)
library(betareg)
library(broom)
library(caret)
library(janitor)
library(performance)
library(gam)
library(mgcv)
library(pls)
theme_set(theme_minimal())
```

Importing data. Veg metrics and percent burned already added in 2023 Burned Unburned Locs Veg .rmd
```{r}
#### Importing data ####
s23rx <- read_xlsx("/Volumes/Samsung_T5/BOBWHITE_DATA/Clean/SummerLocations/Summer2023/s23rx_no6.xlsx") 
```


Centering and scaling veg covariates
```{r}
s23rx_sc <- s23rx %>% 
  mutate_at(vars(hg:ps), .funs = function(x){as.numeric(scale(x, center = T))}) %>%
  mutate(pib = pib + 0.00001)

s23rx_sc1 <- s23rx_sc %>%
  select(pib, dsf, hg:ps)

s23_rx_uniq <- s23rx_sc1[!duplicated(s23rx_sc1), ]
```

Adding in temp data
```{r}
#### importing temp data ####
temp23 <- read_csv("/Volumes/Samsung_T5/BOBWHITE_DATA/Clean/Other/IchTempData.csv") %>%
  clean_names() %>%
  mutate(date = as.Date(date, tryFormats = "%m-%d-%Y")) 

temp23 <- temp23[temp23$date %in% seq(as.Date("2023-03-22"), by="day",length.out=143),]

temp23 <- temp23 %>% 
  select(date, air_temp_max_c)

start_date_23 <- ymd("2023-03-21")

temp23$dsf <- as.numeric(temp23$date - start_date_23)

s23_rx_uniq_temp <- s23_rx_uniq %>%
  left_join(temp23, by = "dsf")


s23_rx_uniq_temp$year <- 2023

s23_rx_uniq_temp_mod <- s23_rx_uniq_temp %>%
  mutate_at(vars(air_temp_max_c), .funs = function(x){as.numeric(scale(x, center = T))})
```

Plotting use over time
```{r}
#### Plotting use over time ####

# If date is not in Date format
# s23rx$date <- as.Date(s23rx$date)

s23_rx_uniq %>%
  ggplot(aes(dsf, pib)) +
  geom_point(size = 2) +
  labs(x = "Days since fire",
       y = "Proportion of bobwhite in burned areas") +
  theme_bw() +
  scale_x_continuous(breaks = c(0, 30, 60, 90, 120, 150), limits = c(NA, 150)) +
  theme_bw() +
  theme(legend.position = "none",
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line("black"),
        axis.ticks = element_line("black"),
        axis.text.x = element_text(colour = "black", size = 12),
        axis.title.y = element_text(colour = "black", size = 12),
        axis.title.x = element_text(colour = "black", size = 12),
        legend.title = element_text(colour = "black", size = 12),
        legend.text = element_text(colour = "black", size = 12),
        axis.text.y = element_text(colour = "black", size = 12),
        plot.margin = margin(t = 5, r = 10, b = 5, l = 5, unit = "pt"))

```

2023: Univariate candidate models (GAM)
```{r}

#### Candidate models (GAM) 2023 #### 

gmodel1 <- gam(pib ~ dsf,
                     data = s23_rx_uniq,
                     family=betar(link = "logit"),
                     na.action = "na.pass")


gmodel2 <- gam(pib ~ pg, 
                  data = s23_rx_uniq,
                  family = betar(link = "logit"),
                  na.action = "na.pass")

gmodel3 <- gam(pib ~ pf, 
                  data = s23_rx_uniq,
                  family = betar(link = "logit"),
                  na.action = "na.pass")

gmodel4 <- gam(pib ~ ps,
                  data = s23_rx_uniq,
                  family = betar(link = "logit"),
                  na.action = "na.pass")

gmodel5 <- gam(pib ~ hg,
                  family = betar(link = "logit"),
                  data = s23_rx_uniq,
                  na.action = "na.pass")


gmodel6 <- gam(pib ~ hf,
                  family = betar(link = "logit"),
                  data = s23_rx_uniq,
                  na.action = "na.pass")

gmodel7 <- gam(pib ~ hs,
                   family = betar(link = "logit"),
                  data = s23_rx_uniq,
                  na.action = "na.pass")

gmodel8 <- gam(pib ~ 1,
                  family = betar(link = "logit"),
                  data = s23_rx_uniq,
                  na.action = "na.pass")

#gmodel9 <- gam(pib ~  dsf + air_temp_max_c,
 #                 family = betar(link = "logit"),
  #                data = s23_rx_uniq_temp_mod,
   #               na.action = "na.pass")



#gmodel10 <- gam(pib ~ air_temp_max_c,
            #      family = betar(link = "logit"),
             #     data = s23_rx_uniq_temp_mod,
              #    na.action = "na.pass")
```

Ranking candidate models (GAM)
```{r}

#### Ranking gam models 2023 ####
# Creating a data frame to store model names and their AIC values
model_names <- c(paste0("gmodel", 1:8))


model_list <- list(gmodel1, gmodel2, gmodel3, gmodel4,
                   gmodel5, gmodel6, gmodel7, gmodel8)


model_aic <- sapply(model_list, AIC)
model_nll <- sapply(model_list, logLik)

# Combine the model names and AIC values into a data frame
model_comparison <- data.frame(Model = model_names, AIC = model_aic, nll = - model_nll)

# Sort the models by AIC value
model_comparison_sorted <- model_comparison %>% 
  arrange(AIC)

# View the sorted data frame
model_comparison_sorted %>% 
  view()

# Calculate the minimum AIC value among all models
min_aic <- min(model_comparison_sorted$AIC)

# Add a column for Delta AIC
model_comparison_sorted <- model_comparison_sorted %>%
  mutate(Delta_AIC = AIC - min_aic)

# View the updated data frame
model_comparison_sorted %>% view()

model_comparison_sorted$rll <- exp(-0.5*model_comparison_sorted$Delta_AIC)
model_comparison_sorted$weight <- model_comparison_sorted$rll/sum(model_comparison_sorted$rll)
model_comparison_sorted$weight <- round(model_comparison_sorted$weight,3)

model_comparison_sorted %>% view

write_xlsx(model_comparison_sorted, "/Users/jeffgrayum/Downloads/model_comp23_No6.xlsx")
```

```{r}
#### Making predict plot ####

dsf_values <- seq(min(s23_rx_uniq$dsf, na.rm = TRUE), 
                 max(s23_rx_uniq$dsf, na.rm = TRUE), 
                 length.out = 1000)

dsf_df <- data.frame(
  hs = rep(mean(s23_rx_uniq$hs), length(dsf_values)),
  hg = rep(mean(s23_rx_uniq$hg), length(dsf_values)),
  hf = rep(mean(s23_rx_uniq$hf), length(dsf_values)),
  pg = rep(mean(s23_rx_uniq$pg), length(dsf_values)),
  pf = rep(mean(s23_rx_uniq$pf), length(dsf_values)),
  ps = rep(mean(s23_rx_uniq$ps), length(dsf_values)),
  dsf = dsf_values
)

dsf_pred <- data.frame(predict(gmodel1, newdata = dsf_df, type = "response", se.fit = TRUE))


dsf_pred$lowCI = (dsf_pred$fit - (1.96 * dsf_pred$se.fit))
dsf_pred$upperCI = (dsf_pred$fit + (1.96 * dsf_pred$se.fit))


dsf_pred %>%
  ggplot(aes(dsf_values, y = fit)) +
  geom_line() +
  geom_ribbon(aes(ymin = lowCI, ymax = upperCI), fill = "red", alpha = 0.2) +
  labs(x = "Days since fire",
       y = "Proportion of bobwhite predicted in burned areas") +
  theme_bw() +
  theme(legend.position = "none",
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line("black"),
        axis.ticks = element_line("black"),
        axis.text.x = element_text(colour = "black", size = 12),
        axis.title.y = element_text(colour = "black", size = 12),
        axis.title.x = element_text(colour = "black", size = 12),
        legend.title = element_text(colour = "black", size = 12),
        legend.text = element_text(colour = "black", size = 12),
        axis.text.y = element_text(colour = "black", size = 12),
        plot.margin = margin(t = 5, r = 10, b = 5, l = 5, unit = "pt"))
```

Let's try combining 2023/23 data
```{r}
s22_rx_uniq <- s22_rx_uniq %>%
  rename(dsf = DSF)

s22_23_rx_unique <- rbind(
  s22_rx_uniq, s23_rx_uniq
)

s22_rx_uniq_temp_mod <- s22_rx_uniq_temp_mod %>%
  rename(dsf = DSF)

s22_23_temp_mod <- rbind(s22_rx_uniq_temp_mod, s23_rx_uniq_temp_mod)
```

Univariate candidate models (GAM) 2022/23
```{r}
#### Candidate models (GAM) 2022/2023 #### 

gmodel1 <- gam(pib ~ dsf,
                     data = s22_23_temp_mod,
                     family=betar(link = "logit"),
                     na.action = "na.pass")


gmodel2 <- gam(pib ~ pg, 
                  data = s22_23_temp_mod,
                  family = betar(link = "logit"),
                  na.action = "na.pass")

gmodel3 <- gam(pib ~ pf, 
                  data = s22_23_temp_mod,
                  family = betar(link = "logit"),
                  na.action = "na.pass")

gmodel4 <- gam(pib ~ ps,
                  data = s22_23_temp_mod,
                  family = betar(link = "logit"),
                  na.action = "na.pass")

gmodel5 <- gam(pib ~ hg,
                  family = betar(link = "logit"),
                  data = s22_23_temp_mod,
                  na.action = "na.pass")


gmodel6 <- gam(pib ~ hf,
                  family = betar(link = "logit"),
                  data = s22_23_temp_mod,
                  na.action = "na.pass")


gmodel7 <- gam(pib ~ hs,
                   family = betar(link = "logit"),
                  data = s22_23_temp_mod,
                  na.action = "na.pass")

gmodel8 <- gam(pib ~ 1,
                  family = betar(link = "logit"),
                  data = s22_23_temp_mod,
                  na.action = "na.pass")

#gmodel9 <- gam(pib ~  dsf + air_temp_max_c,
                #  family = betar(link = "logit"),
               #   data = s22_23_temp_mod,
               #   na.action = "na.pass")



#gmodel10 <- gam(pib ~ air_temp_max_c,
                 # family = betar(link = "logit"),
                #  data = s22_23_temp_mod,
                #  na.action = "na.pass")

```

Ranking candidate models (gam)
```{r}

#### Ranking gam models 2022/23 ####
# Creating a data frame to store model names and their AIC values
model_names <- c(paste0("gmodel", 1:8))


model_list <- list(gmodel1, gmodel2, gmodel3, gmodel4,
                   gmodel5, gmodel6, gmodel7, gmodel8)


model_aic <- sapply(model_list, AIC)
model_nll <- sapply(model_list, logLik)

# Combine the model names and AIC values into a data frame
model_comparison <- data.frame(Model = model_names, AIC = model_aic, nll = - model_nll)

# Sort the models by AIC value
model_comparison_sorted <- model_comparison %>% 
  arrange(AIC)

# View the sorted data frame
model_comparison_sorted %>% 
  view()

# Calculate the minimum AIC value among all models
min_aic <- min(model_comparison_sorted$AIC)

# Add a column for Delta AIC
model_comparison_sorted <- model_comparison_sorted %>%
  mutate(Delta_AIC = AIC - min_aic)

# View the updated data frame
model_comparison_sorted %>% view()

model_comparison_sorted$rll <- exp(-0.5*model_comparison_sorted$Delta_AIC)
model_comparison_sorted$weight <- model_comparison_sorted$rll/sum(model_comparison_sorted$rll)
model_comparison_sorted$weight <- round(model_comparison_sorted$weight,3)

model_comparison_sorted %>% view

write_xlsx(model_comparison_sorted, "/Users/jeffgrayum/Downloads/model_comparison22_23fixed.xlsx")
```




Let's try seeing if we can predict prob of burned/unburned by using distance to edge.
```{r}
gmodel_dte <- gam(burned ~ distance_to_edge,
                 family = betar(link = "logit"),
                 data = s23rx_sc,
                 na.action = "na.pass")

dte_values <- seq(min(s23rx_sc$distance_to_edge, na.rm = TRUE), 
                 max(s23rx_sc$distance_to_edge, na.rm = TRUE), 
                 length.out = 1000)

dte_df <- data.frame(
  distance_to_edge = dte_values
)

dte_pred <- data.frame(predict(gmodel_dte, newdata = dte_df, type = "response", se.fit = TRUE))


dte_pred$lowCI = (dte_pred$fit - (1.96 * dte_pred$se.fit))
dte_pred$upperCI = (dte_pred$fit + (1.96 * dte_pred$se.fit))

dte_pred %>%
  ggplot(aes(dte_values, fit)) +
  geom_line() +
  geom_ribbon(aes(ymin = lowCI, ymax = upperCI), fill = "blue", alpha = 0.1) +
  labs(x = "Distance to edge (m) of burn compartment",
       y = "Probability of location falling in burned area",
       title = "Predicted probability of bobwhite being in burned area based on distance to edge") +
  scale_y_continuous(labels = percent)
```


Now, let's see if we can predict probability of use based on distance to edge of burned compartments. We will need to important real/random locations from Ch1 analysis.
```{r}
#### Adding dte to real/random points ####

#Importing map of burned areas, summer 2023.
#Reading burn compartments in again, but as an sf object
burn_compartments_2023 <- st_read("/Volumes/Samsung_T5/BOBWHITE_DATA/Maps/2023_burn_compartments/burn 2023 map/NoAG_NoHW/2023_burn_units_NoAg_NoHW.shp")

# Converting burn_compartments_2022 polygons to line geometries that represent their boundaries. This is necessary to calc min distances later.
burn_compartment_boundaries <- st_boundary(burn_compartments_2023)

# Converting tibble of points to an sf object
s23_locs_sf <- st_as_sf(rsfData_modified_summer2023, coords = c("x_", "y_"), crs = st_crs(burn_compartments_2023))

# Calculating the distance from each point to the nearest line boundary of the burn compartments, storing it in matrix
distance_matrix <- st_distance(s23_locs_sf, burn_compartment_boundaries)

# Applying the function min() to each row (1 does this -- we'd use 2 for columns) in the distance matrix
min_distances <- apply(distance_matrix, 1, min)

# Add the minimum distance to the original tibble as a new column
rsfData_modified_summer2023$distance_to_edge <- min_distances


#Now, we can predict real/random
gmodel_prob_use <- gam(case ~ distance_to_edge,
                 family = betar(link = "logit"),
                 data = rsfData_modified_summer2023,
                 na.action = "na.pass")  
  
  
dte_values <- seq(min(rsfData_modified_summer2023$distance_to_edge, na.rm = TRUE), 
                 max(rsfData_modified_summer2023$distance_to_edge, na.rm = TRUE), 
                 length.out = 1000)

dte_df <- data.frame(
  distance_to_edge = dte_values
)

prob_use_pred <- data.frame(predict(gmodel_prob_use, newdata = dte_df, type = "response", se.fit = TRUE))


prob_use_pred$lowCI = (prob_use_pred$fit - (1.96 * prob_use_pred$se.fit))
prob_use_pred$upperCI = (prob_use_pred$fit + (1.96 * prob_use_pred$se.fit))

prob_use_pred %>%
  ggplot(aes(dte_values, fit)) +
  geom_line() +
  geom_ribbon(aes(ymin = lowCI, ymax = upperCI), fill = "blue", alpha = 0.1) +
  labs(x = "Distance to burn edge (m)",
       y = "Probability of use",
       title = "Predicted probability of use based on distance to edge") +
  scale_y_continuous(labels = percent)

```


```{r}
#### Plotting veg data vs percent in burned ####
s23rx_piv <- s23rx %>%
  select(pg:hs, pib) %>%
  pivot_longer(pg:hs, names_to = "Measurement", values_to = "Value")


s23rx_piv %>% 
  dplyr::filter(Measurement %in% c("pg", "pf", "ps")) %>%
  mutate(Measurement = factor(Measurement, levels = c("pg", "pf", "ps"))) %>%
  ggplot(aes(Value, pib)) +
  geom_point(color = "Midnight Blue") +
  facet_wrap(~ Measurement, scales = "free") +
  labs(x = "% cover",
       y = "% of birds in burned area",
       title = "2023: Percent cover of grasses, forbs, and shrubs vs percent of birds in burned areas") +
  theme_bw()

s23rx_piv %>% 
  dplyr::filter(Measurement %in% c("hg", "hf", "hs")) %>%
  mutate(Measurement = factor(Measurement, levels = c("hg", "hf", "hs"))) %>%
  ggplot(aes(Value, pib)) +
  geom_point(color = "Midnight Blue") +
  facet_wrap(~ Measurement, scales = "free") +
  labs(x = "Height (cm)",
       y = "% of birds in burned area",
       title = "2023: Height of grasses, forbs and shrubs vs percent of birds in burned areas") +
  theme_bw()


s23rx %>%
  ggplot(aes(pg, pib)) +
  geom_point(color = "Midnight Blue") +
  labs(x = "% cover of grasses",
       y = "Perenct of birds in burned area",
       title = "Percent cover grasses vs Percent of birds in burned areas") +
  theme_bw()

s23rx %>%
  ggplot(aes(pf, pib)) +
  geom_point(color = "Midnight Blue") +
  labs(x = "% cover of forbs",
       y = "Perenct of birds in burned area",
       title = "Percent cover forbs vs Percent of birds in burned areas") +
  theme_bw()


s23rx %>%
  ggplot(aes(ps, pib)) +
  geom_point(color = "Midnight Blue") +
  labs(x = "% cover of shrubs",
       y = "Perenct of birds in burned area",
       title = "Percent cover shrubs vs Percent of birds in burned areas")

s23rx %>%
  ggplot(aes(hg, pib)) +
  geom_point(color = "Midnight Blue") +
  labs(x = "Height of grasses (cm)",
       y = "Perenct of birds in burned area",
       title = "Height of grasses vs Percent of birds in burned areas")

s23rx %>%
  ggplot(aes(hf, pib)) +
  geom_point(color = "Midnight Blue") +
  labs(x = "Height of forbs (cm)",
       y = "Perenct of birds in burned area",
       title = "Height of forbs vs Percent of birds in burned areas")

s23rx %>%
  ggplot(aes(hs, pib)) +
  geom_point(color = "Midnight Blue") +
  labs(x = "Height of shrubs (cm)",
       y = "Perenct of birds in burned area",
       title = "Height of shrubs vs Percent of birds in burned areas")
  
```



```{r}

degree_label <- paste0("Max ambient temp (", "\u00B0", "F)")

# Color mapping
color_map <- setNames(c("red", "black", "blue"), c(degree_label, "% Bobwhite in burned units", "Bobwhite thermal stress threshold"))

# Assuming s22rx exists and has the relevant columns
# Create dashed line data
dashed_line_data <- data.frame(dsf = c(min(s23rx$dsf), max(s23rx$dsf)),
                               y = c(86.5 / 2, 86.5 / 2))

s23_rx_uniq_temp %>%
  ggplot(aes(x = dsf)) +
  geom_line(aes(y = (air_temp_max_c * 1.8 + 32) / 2, color = degree_label)) +
  geom_point(aes(y = pib * 100, color = "% Bobwhite in burned units")) +
  geom_line(data = dashed_line_data, aes(y = y, color = "Bobwhite thermal stress threshold"), linetype = "dashed") +
  scale_y_continuous(
    name = "% Bobwhite in burned units",
    sec.axis = sec_axis(~ . * 2, name = degree_label)
  ) +
  scale_color_manual(
    values = color_map,
    name = ""
  ) +
  labs(x = "Days since fire",
       y = "% sampled bobwhite in burned units",
       title = "Use of burned units after prescribed fire\nand max ambient temp (2023)") +
  theme_bw() +
  scale_x_continuous(breaks = c(0, 30, 60, 90, 120))


#* 1.8 + 32) / 2
```


```{r}
s23dist <- s23rx %>%
  select(burned, distance_to_edge)

s22_s23_dist <- rbind(s22dist, s23dist)

s22_s23_dist %>%
  rename(Location = burned) %>%
  mutate(Location = factor(Location, labels = c("Unburned", "Burned"))) %>%
  ggplot(aes(x = distance_to_edge, fill = Location)) +
  geom_histogram(alpha = 0.8) +
  labs(x = "Distance to edge of burned unit (m)",
       y = "Count",
       title = "Distance of bobwhite locations to edge of burned units") +
  theme_bw()

+
  scale_x_continuous(breaks = c(0, 100, 200, 300, 400, 500, 600)) +
  theme(panel.grid.minor = element_blank())

s22_s23_dist %>%
  dplyr::filter(burned == 0) %>%
  ggplot(aes(x = distance_to_edge)) +
  geom_histogram()
```

```{r}
global_model <- gam(pib ~ pg + pf + ps + hg + hf + hs + dsf,
 family = betar(link = "logit"),
 data = s22_23_rx_unique,
 na.action  = "na.pass")

summary(global_model)
```


```{r}
s23temp <- s23temp %>%
  rename(dsf = DSF)

s23_rx_uniq_temp <- s23_rx_uniq %>%
  left_join(s23temp, by = join_by(dsf), relationship = "many-to-many")

s23_rx_uniq_temp_mod <- s23_rx_uniq_temp %>%
  mutate_at(vars(air_temp_max_c), .funs = function(x){as.numeric(scale(x, center = T))})


gmodel9 <- gam(pib ~  dsf + air_temp_max_c,
                  family = betar(link = "logit"),
                  data = s23_rx_uniq_temp_mod,
                  na.action = "na.pass")

gmodel10 <- gam(pib ~ air_temp_max_c,
                  family = betar(link = "logit"),
                  data = s23_rx_uniq_temp_mod,
                  na.action = "na.pass")



s22_23_rx_temp_unique

```

```{r}
#### Testing distance to edge and DSF #### 

s23rx_bu <- s23rx %>%
  dplyr::filter(burned == 1)

dist_model23 <- lmer(distance_to_edge ~ dsf + (1 | band_numb),
                   data = s23rx_bu)

summary(dist_model23)
```

